---
title: "Untitled Draft"
author: "Report Author"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
drake::loadd(all_data)
drake::loadd(variant_step_plot)
drake::loadd(type_dict)
library(tidyverse)
```

# Data

```{r}
all_data
```

# Number of unique reads by number of reads

```{r}
variant_step_plot +
  theme_minimal() +
  labs(
    title = "One color for each sample",
    x = "Number of sequenced alleles",
    y = "Number of unique alleles"
    )
```

# Count of alleles in 1_M.bscflags sample

```{r}
n_obs_1_M <- all_data %>%
      filter(sample == "1_M.bscflags") %>%
  nrow()

all_data %>%
      filter(sample == "1_M.bscflags") %>%
  count(htype, sort = TRUE) %>%
  ggplot(aes(n, htype)) +
  geom_col() +
  geom_vline(xintercept = n_obs_1_M * c(0.01, 0.001), color = c("red", "blue")) +
  theme_minimal() +
  scale_x_log10() +
  labs(title = "Notice that x-axis is on log-scale",
       subtitle = "Proportion: blue = 0.001, red = 0.01")
```

# Proportion 0.05

```{r}
drake::loadd(prop_0.05)
prop_0.05
```

# Proportion 0.02

```{r}
drake::loadd(prop_0.02)
prop_0.02
```

# Proportion 0.01

```{r}
drake::loadd(prop_0.01)
prop_0.01
```


# Proportion 0.001

```{r}
drake::loadd(prop_0.001)
prop_0.001
```

# Clustering

We calculated the number of patterns that appears at least 1% of the times for each sample and amplicon. This gives us data on the following form:

```{r}
drake::loadd(amplicon_0.01)
amplicon_0.01 %>% rename(n_pattern = n_dist)
```

We then transform the data to have amplicons as columns and samples as rows. Do mean-imputation of missing values. Then we project the data down to a lower dimensional space and see if we can find cluster with an kmeans algorithm.

```{r}
drake::loadd(amplicon_pca_clusters)

amplicon_pca_clusters <- amplicon_pca_clusters %>% 
  mutate(.cluster = factor(.cluster, c(2, 1, 3), c("red", "green", "blue")))

amplicon_pca_clusters %>%
  ggplot(aes(PC1, PC2, color = .cluster)) +
  geom_point() +
  theme_minimal() +
  labs(title = "PCA projection")
```

```{r}
drake::loadd(amplicon_umap_clusters)

amplicon_umap_clusters <- amplicon_umap_clusters %>%
  mutate(.cluster = factor(.cluster, c(1, 2, 3), c("red", "green", "blue")))

amplicon_umap_clusters %>%
  ggplot(aes(umap_1, umap_2, color = .cluster)) +
  geom_point() +
  theme_minimal() +
  labs(title = "UMAP projection")
```

# Clusters applied to first chart

We take the clusters we found in the UMAP project and color the first chart accordingly.

```{r}
drake::loadd(variant_counts)
variant_counts %>%
  left_join(amplicon_umap_clusters, by = "sample") %>%
  ggplot(aes(readseq, n_dis, group = sample, color = .cluster)) +
  geom_step() +
  guides(color = "none") +
  theme_minimal() +
  labs(
    title = "Colored accoring to UMAP clusters",
    x = "Number of sequenced alleles",
    y = "Number of unique alleles"
    )
```

We take the clusters we found in the PCA project and color the first chart accordingly.

```{r}
drake::loadd(variant_counts)
variant_counts %>%
  left_join(amplicon_pca_clusters, by = "sample") %>%
  ggplot(aes(readseq, n_dis, group = sample, color = .cluster)) +
  geom_step() +
  guides(color = "none") +
  theme_minimal() +
  labs(
    title = "Colored accoring to PCA clusters",
    x = "Number of sequenced alleles",
    y = "Number of unique alleles"
    )
```

# Clusters

```{r}
left_join(
  amplicon_pca_clusters %>% select(sample, pca_cluster = .cluster),
  amplicon_umap_clusters %>% select(sample, umap_cluster = .cluster)
) %>%
  left_join(type_dict) %>%
  arrange(umap_cluster) %>%
  knitr::kable()
```

## Reproducibility

<details><summary>Reproducibility receipt</summary>

```{r}
## datetime
Sys.time()

## session info
sessionInfo()
```

</details>
